Stability of variational quantum dynamics with Gaussian approximants

### 2013-07-16 paper.tex
writing

We need a way to discretise the function $g$.  There are two usual ways to do this: sampling on a grid, or taking components along a fixed basis.  The size of state space rules these out, so we must sample $g$ at a set of irregularly spaced amplitudes $A=\{α₁,…,α_R\}$, corresponding to a row of kets $|A〉=\bigl(|α₁〉\ ⋯\ |α_R〉\bigr).  Equivalently, we take components of $f$ in function space along the set of vectors $χ_r(α)=w_rδ(α-α_r)$.  Here, $w_r$ is a real weight, intended to compensate for the irregularity of the sample set, so that the sampled vector $\vec g=\bigl(w₁g(α₁),…,w_Rg(α_R)\bigr)^{\rm T}$ has a similar norm to the function $g$.

@@

Putting these conditions together, with regularisation parameters $λ$ and $ε$, gives the problem
$$\pmatrix{|Dψ(z)〉& 0 \cr λI & 0 \cr Dσ_R(z) & Cσ_R(z)}
	\pmatrix{\dot z\cr \dot z*} =
	\pmatrix{H|ψ(z)〉\cr 0\cr |σ_R||\dot z|_{\rm opt}}.$$
This raises a technical problem, because we need to constrain $\dot z*$ to be the conjugate of $\dot z$.  Alternatively, if we write $z=x+iy$, we need to solve a complex least squares problem under the constraint that the solution is real.  Neither of those sounds trivial, but it shouldn't be impossible.

>>>>

We need a way to discretise the function $g$.  There are two usual ways to do this: sampling on a grid, or taking components along a fixed basis.  The size of state space rules these out, so we must sample $g$ at a set of irregularly spaced amplitudes $A=\{α₁,…,α_R\}$, corresponding to a row of kets $|A〉=\bigl(|α₁〉\ ⋯\ |α_R〉\bigr)$.  Equivalently, we take components of $f$ in function space along the set of vectors $χ_r(α)=w_rδ(α-α_r)$.  Here, $w_r$ is a real weight, intended to compensate for the irregularity of the sample set, so that the sampled vector $\vec g=\bigl(w₁g(α₁),…,w_Rg(α_R)\bigr)^{\rm T}$ has a similar norm to the function $g$.

@@

Putting these conditions together, with regularisation parameters $λ$ and $ε$, gives the least squares problem
$$\pmatrix{|Dψ(z)〉& 0 \cr λI & 0 \cr Dσ_R(z) & Cσ_R(z)}
	\pmatrix{\dot z\cr \dot z*} ≈
	\pmatrix{H|ψ(z)〉\cr 0\cr |σ_R||\dot z|_{\rm opt}}.$$
This raises a technical problem, because we need to constrain $\dot z*$ to be the conjugate of $\dot z$.  In the space of complex numbers, this constraint is not linear, because $(cz)*≠cz*$.  However, that is true if $c$.  Moreover, if $z=x+iy$, then $|z-w|²=|x-\Re w|²+|y-\Im w|²$, so a least squares problem in $z$ has the same solution as the equivalent least squares problem in $x$ and $y$.

Consider the problem 
$$ \pmatrix{A & B}\pmatrix{z\cr z*}≈c,$$
with $z∈C^m$, $c∈C^n$, and $A,B∈C^{m×n}$.  If $z=x+iy$, the left hand side can be rearranged
$$\pmatrix{A & B}\pmatrix{z\cr z*}
	=\pmatrix{A & B}\pmatrix{I&iI\cr I&-iI}\pmatrix{x\cr y}
	=\pmatrix{A+B & iA-iB}\pmatrix{x\cr y}.$$
Separating the rows into real and imaginary parts gives
$$\pmatrix{\Re (A+B) & \Re(iA-iB)\cr\Im(A+B) & \Im(iA-iB)}\pmatrix{x\cr y}
	=\pmatrix{\Re A+\Re B & \Im B-\Im A\cr\Im A+\Im B & \Re A-\Re B)}
		\pmatrix{x\cr y}≈\pmatrix{\Re c\cr \Im c}.$$
